---
title: "R1 - Data analyis"
author: "Your name"
date: "`r Sys.time()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r message = F}
library(tidyverse)
library(lubridate)
```

## a.
Download  data Atlanta [weather](http://www.richardtwatson.com/data/ATLweather.csv) (a timestamp, air temperature, humidity, and precipitation), and  [electricity prices](http://www.richardtwatson.com/data/electricityprices.csv) (a timestamp and cost in cents per kWh).
Merge the files for electricity price, and weather.

```{r message = F}
# R code for a.
```

## b.
Compute the correlation between temperature and electricity price. What do you conclude?

```{r}
# R code for b.
```

### Conclusion


## c.
Extract the data for August and redo the correlation What do you conclude?

```{r}
# R code for c.
```
### Conclusion

## d.
Extract the data for January and redo the correlation What do you conclude?

```{r}
# R code for d.
```

### Conclusion

## e.
Download data for solar radiation (a timestamp and solar radiation in watts/m2) for Athens. Compute the average, min, and max for solar radiation for the year (one value for each of the three measures). 

### Average, min, and max

```{r}
# R code for e.
```

## f. 
Assuming the total area for capturing solar energy by PV cells for a house is 25m2 (269 square feet)) and solar cells are 20% efficient (i.e.., 20% of the photons received are converted into electrons). How much electricity in kWh will be generated on average each day in August and January?

```{r}
# R code for f.
```

### Findings


