---
title: "R1 - Data reporting and analysis"
author: "Richard T. Watson"
date: "`r Sys.time()`"
output:
  pdf_document: default
  html_document: default
---

```{r message = F}
library(tidyverse)
library(lubridate)
```

## a.
Access the data Atlanta [weather](http://www.richardtwatson.com/data/ATLweather.csv) (a timestamp, air temperature, humidity, and precipitation), and  [electricity prices](http://www.richardtwatson.com/data/electricityprices.csv) (a timestamp and cost in cents per kWh).
Merge the files for electricity price, and weather.

```{r message = F}
url <-  "http://www.richardtwatson.com/data/ATLweather.csv"
w <- read_delim(url,delim=',')
url <-  "http://www.richardtwatson.com/data/electricityprices.csv"
e <- read_delim(url,delim=',')
m <- inner_join(w,e,by=c("Timestamp" = 'timestamp'))
```

## b.
Compute the correlation between temperature and electricity price. What do you conclude?

```{r}
cor.test(m$Temperature,m$cost)
```

### Conclusion
* As the p-value is less than .05, conclude that there is a relationship between temperature and electricity cost. As the temperature increases, electricity prices increase.
* The correlation is *small* (0.18) as it is between .1 and .3, as shown in the following [table](https://en.wikipedia.org/wiki/Effect_size):

| Correlation coefficient | Effect size | 
|:-----------------------:|:-----------:|
| .10 - .30               |    Small    |
| .30 - .50               |   Moderate  |
| > .50                   |    Large    |

* Given the small correlation, there are likely other factors that influence electricity cost, such as generation capacity

**Note:** The p-value tells you whether there is a relationship and the correlation coefficient indicates the size of that relationship.

## c. Extract the data for August and redo the correlation What do you conclude?

```{r}
Aug <- m %>% filter((month(Timestamp) == 8))
cor.test(Aug$Temperature,Aug$cost)
```
### Conclusion
* As the p-value is less than .05, conclude that there is a relationship between temperature and electricity cost during August
* The correlation is  *large* (0.56).
There appears to be generation capacity problem in August, usually the hottest months of the year, so electricity prices are raised to dampened demand on hot days.

## d.
Extract the data for January and redo the correlation What do you conclude?

```{r}
Jan <- m %>% filter((month(Timestamp) == 1))
cor.test(Jan$Temperature,Jan$cost)
```
### Conclusion
* As the p-value is less than .05, conclude that there is a relationship between temperature and electricity cost during winter
* The correlation is negative and *moderate* (-.4).
There appears to be generation capacity problem in January, usually the coldest month of the year, so electricity prices are raised to dampened demand on cold days.

## e.
Download the data for [solar radiation](http://www.richardtwatson.com/data/SolarRadiationAthens.csv) (a timestamp and solar radiation in watts/m2) for Athens. Compute the average, min, and max for solar radiation for the year (one value for each of the three measures). 

### Average, min, and max

```{r}
url <- "http://www.richardtwatson.com/data/SolarRadiationAthens.csv"
s <- read_delim(url,delim=',')
mean(s$SolarWatt)
min(s$SolarWatt)
max(s$SolarWatt)

```

## f. 
Assuming the total area for capturing solar energy by PV cells for a house is 25m2 (269 square feet)) and solar cells are 20% efficient (i.e., 20% of the radiation received is converted into electricity). How much electricity in kWh will be generated on average each day in August and January?

### Explanation
* A solar panel of 20% efficiency receiving 200 Watts/m^2 of solar radiation will generate 40 Watts/m^2 of electricity
* A solar panel system generating 1000 Watts consistently for one hour generates 1 kWh. In a day it would generate 24kWh (1000*24/1000). Of course, the sun does not shine consistently for 24 hours.


```{r}
Aug <- s %>% filter((month(TimeStamp) == 8))
round(mean(Aug$SolarWatt)*24/1000*25*.2,2)
Jan <- s %>% filter((month(TimeStamp) == 1))
round(mean(Jan$SolarWatt)*24/1000*25*.2,2)
```

### Findings
About 24.56 kWh will be generated each day in August.
About 14.43 kWh will be generated each day in January.

