---
title: "R1 report"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

## Author: Richard T. Watson
```{r echo = FALSE, results = 'asis'}
cat(paste('\n###','Date:', Sys.Date()))
```

```{r message = F}
library(dplyr)
library(readr)
library(lubridate)
```

## a.
Download  data Atlanta [weather](http://www.richardtwatson.com/data/ATLweather.csv) (a timestamp, air temperature, humidity, and precipitation), and  [electricity prices](http://www.richardtwatson.com/data/electricityprices.csv) (a timestamp and cost in cents per kWh).
Merge the files for electricity, price, and weather.

```{r}
url <-  "http://www.richardtwatson.com/data/ATLweather.csv"
w <- read_delim(url,delim=',')
url <-  "http://www.richardtwatson.com/data/electricityprices.csv"
e <- read_delim(url,delim=',')
m <- inner_join(w,e,by=c("Timestamp" = 'timestamp'))
```

## b.
Compute the correlation between temperature and electricity price. What do you conclude?

```{r}
cor.test(m$Temperature,m$cost)
```

### Conclusion
* As the p-value is less than .05, conclude that there is a relationship between temperature and electricity cost. As the temperature increases, electricity prices increase.
* The correlation is *small* (0.18) as it is between .1 and .3, as shown in the following [table](https://en.wikipedia.org/wiki/Effect_size):

| Correlation coefficient | Effect size | 
|-------------------------|:-----------:|
| .10 - .30               |    Small    |
| .30 - .50               |   Moderate  |
| > .50                   |    Large    |

* Given the small correlation, there are likely other factors that influence electricity cost, such as generation capacity

**Note:** The p-value tells you whether there is a relationship and the correlation coefficient indicates the size of that relationship.

## c.
Extract the data for July through September (Summer) and redo the correlation What do you conclude?

```{r}
Summer <- m %>% filter((month(Timestamp) >= 7 & month(Timestamp) <= 9))
cor.test(Summer$Temperature,Summer$cost)
```
### Conclusion
* As the p-value is less than .05, conclude that there is a relationship between temperature and electricity cost during summer.
* The correlation is  *large* (0.51).
There appears to be generation capacity problem in summer, usually the hottest months of the year, so electricity prices are raised to dampened demand on hot days.

## d.
Extract the data for January through March (Winter) and redo the correlation What do you conclude?

```{r}
Winter <- m %>% filter((month(Timestamp) >= 1 & month(Timestamp) <= 3))
cor.test(Winter$Temperature,Winter$cost)
```
### Conclusion
* As the p-value is less than .05, conclude that there is a relationship between temperature and electricity cost during winter
* The correlation is negative and *moderate* (-.32).
There appears to be generation capacity problem in winter, usually the coldest months of the year, so electricity prices are raised to dampened demand on cold days.

## f.
Download the Athens data for [solar](http://www.richardtwatson.com/data/SolarRadiationAthens.csv) (a timestamp and solar radiation in watts/m2), Using the Athens solar radiation data, compute the average (one value), min and max for solar radiation. 

### Average, min, and max

```{r}
url <- "http://www.richardtwatson.com/data/SolarRadiationAthens.csv"
s <- read_delim(url,delim=',')
mean(s$SolarWatt)
min(s$SolarWatt)
max(s$SolarWatt)

```

### Monthly average

```{r}
s %>% 
  group_by(month(TimeStamp)) %>% 
  summarize(meanMonth = mean(SolarWatt)
```

## e. 
Assuming the total area for capturing solar energy by PV cells is 25m2 and panels are 20% efficient. How much electricity will be generated in a day?

```{r}
round(mean(s$SolarWatt)*24/1000*25*.2,2)
```

### Findings
About 23kWh will be generated each day. More in summer than winter.

