---
title: "R2 - Visualization"
output:
  pdf_document: default
  html_notebook: default
---

## Author: Richard T. Watson
```{r echo = FALSE, results = 'asis'}
cat(paste('\n###','Date:', Sys.Date()))
```

```{r message = F}
library(readr)
library(ggplot2)
library(dplyr)
library(lubridate)
```

## Prepare the data
```{r}
url <-  "http://www.richardtwatson.com/data/ATLweather.csv"
w <- read_delim(url,delim=',')
url <-  "http://www.richardtwatson.com/data/electricityprices.csv"
e <- read_delim(url,delim=',')
m <- inner_join(w,e,by=c("Timestamp" = 'timestamp'))
```


## a.
Graph the relationship between temperature and electricity price with a smoother.

```{r}
ggplot(m,aes(Temperature,cost)) +
  geom_point(color='skyblue') + 
  geom_smooth() + 
  xlab('Temperature') +
  ylab('Electricity cost in cents/kWh')
```

## b.
Graph the relationship between the temperature and electricity price for winter with a smoother.

```{r}
period <- m %>% filter(month(Timestamp) >= 1 & month(Timestamp) <= 3)
ggplot(period,aes(Temperature,cost)) +
  geom_point(color='skyblue') + 
  geom_smooth() + 
  xlab('Temperature') +
  ylab('Electricity cost in cents/kWh')
```

## c.
Graph the relationship between the temperature and electricity price for summer with a smoother.

```{r}
period <- m %>% filter(month(Timestamp) >= 7 & month(Timestamp) <= 9)
ggplot(period,aes(Temperature,cost)) +
  geom_point(color='skyblue') + 
  geom_smooth() + 
  xlab('Temperature') +
  ylab('Electricity cost in cents/kWh')
```

How do you explain the anomaly in summer prices? Create a graph that supports your explanation.

```{r}
period<- m %>% filter(month(Timestamp) >= 7 & month(Timestamp) <= 9 & wday(Timestamp) > 1 & wday(Timestamp) < 7)
ggplot(period,aes(Temperature,cost)) +
  geom_point(color='skyblue') + 
  geom_smooth() + 
  xlab('Temperature') +
  ylab('Electricity cost in cents/kWh')
```

There is less electricity used on the weekends (days 1 and 7). The preceding graph with these values omitted shows the as temperature increases, prices increase.

## d.
Create a column (geom_col) chart showing average solar radiation for each month in kWh. Label the months 1 through 12 consecutively.
```{r}
url <-  "http://www.richardtwatson.com/data/SolarRadiationAthens.csv"
s <- read_delim(url,delim=',')
s2 <- s %>% 
  group_by(month = month(TimeStamp)) %>% 
  summarize(MeanSolarWatt = mean(SolarWatt*24/1000))
ggplot(s2,aes(month,MeanSolarWatt)) +
  geom_col(fill='orange') + 
  xlab('Month') +
  ylab('Mean solar radiation in kWh/m2/day') +
  scale_x_continuous(breaks = seq(1, 12, by = 1))
```



