---
title: "R2"
author: "Scott Hall and Cindy Zhang"
date: "2/24/2020"
output:
  html_document: null
  pdf_document: default
---

```{r message = F}
library(dplyr)
library(readr)
library(lubridate)
library(ggplot2)
```

```{r}
Athens <- read_csv("http://www.richardtwatson.com/data/SolarRadiationAthens.csv")
electricityprices <- read_csv("http://www.richardtwatson.com/data/electricityprices.csv")
ATLweather <- read_csv("http://www.richardtwatson.com/data/ATLweather.csv")
weather_prices <- inner_join(ATLweather, electricityprices,by=c("Timestamp" = 'timestamp'))
```

## A Graph the relationship between temperature and electricity price with a smoother.

```{r}
ggplot(weather_prices, aes(x=Temperature, y= cost)) + 
        geom_point() +
        geom_smooth()
```
#### Conclusion for part A:
Energy Prices and Tempeture have a parabolic relationship seen by the fact that energy cost is highest when tempeture is at its exteremities.

## B) Graph the relationship between the temperature and electricity price for winter with a smoother. 

```{r}
winter <- weather_prices %>% filter((month(Timestamp) >= 1 & month(Timestamp) <= 3))
ggplot(winter, aes(x=Temperature, y= cost)) + 
        geom_point() +
        geom_smooth()

```
#### Conclusion for part B:
In the winter months, energy prices remain quite low until the temperature drops below 30 degrees.

## C) Graph the relationship between the temperature and electricity price for summer with a smoother. How do you explain the anomaly in summer prices? Create a graph that supports your explanation.